Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-14 paper/paper.tex
Cost of TAlg

\section{Recovering a Perfect Partition}	\label{sec:ppl}
%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
%\maciek{ every algorithm must colocate every communicating pair.
%	\mahmoud{updated!}}
We  focus on a restricted variant of  \OBRP{},
where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
at the beginning and stays there permanently.
We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
This implies that on a request between any two non-collocated nodes,
any online algorithm must eventually collocate them,
as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.
Hence,
any competitive algorithm,
must at some point collocate all nodes of a  connected sub-graph of the communication graph.
We refer to this connected sub-graph as a \emph{component}.
An algorithm that always keeps a component collocated is \emph{component-respecting}.
%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}

@@

In this section we show a lower bound of $\Omega(k \cdot \ell)$ for \OBRP{}.
The constructed input sequence allows OPT to move to a perfect partition, and hence
the construction also constitutes a lower bound for \PPOBRP{}.
This complements an upper bound of $O(k \cdot \ell)$
for \PPOBRP{} that we present in Section \ref{sec:ppl}.

@@

  We say that \TAlg is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.

  \begin{lemma}
    \label{lem:1req}
    Serving each request incurs the cost at most $3\cdot\alpha$ for \TAlg.
  \end{lemma}

  \begin{proof}
    If no partition of \TAlg's components exist, then it performs no repartition, and it pays nothing.
    Hence, in the following we bound the cost of moving to component-respecting partition.

@@

  \begin{theorem}
    \TAlg is $O(\ell)$-competitive.
  \end{theorem}
  \begin{proof}
    Consider a request $(u,v)$.
    If $u$ and $v$ are collocated in the same server, then \TAlg pays nothing.

    \TAlg collocates nodes belonging to one component, hence during a single phase, we have at most $2\cdot \ell$ requests to non-collocated nodes.
    We use Lemma~\ref{lem:1req} to bound the cost of each such request, and therefore \TAlg pays $6\cdot\alpha\cdot \ell$ in a single phase.
    On the other hand, during each phase, OPT pays at least $1$.
  \end{proof}

>>>>

In this section we show a lower bound of $\Omega(k \cdot \ell)$.
The constructed input sequence allows OPT to move to a perfect partition, and hence
the construction also constitutes a lower bound for \PPOBRP{}.
This complements an upper bound of $O(k \cdot \ell)$
for \PPOBRP{} that we present in Section \ref{sec:ppl}.

\section{Recovering a Perfect Partition}	\label{sec:ppl}
%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
\maciek{ every algorithm must colocate every communicating pair.
	\mahmoud{updated!}}
We  focus on a restricted variant of   \PPOBRP{}
where the communication graph can be partitioned into $\ell$ clusters without any inter-cluster edges.
Moreover, we assume an optimal offline algorithm (OPT) moves to this \emph{perfect partition}
at the beginning and stays there permanently.
Then the task of an online algorithm is to "recover" (or learn) the perfect partition while not paying too much more than OPT.
This implies that on a request between any two non-collocated nodes,
any online algorithm must eventually collocate them,
as otherwise it not competitive under a sequence that repeats the same request arbitrarily many times.

\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}

@@

@@

  We say that \TAlg is \emph{component respecting}, i.e., it maintaints the invariant that nodes of each component are located at the same server.
  Observe that this implies that \TAlg never increases any counter above $\alpha$.

  \begin{lemma}
    \label{lem:1req}
    A single repartition incurs the cost at most $3\cdot\alpha$ for \TAlg.
  \end{lemma}

  \begin{proof}
    If no partition of \TAlg's components exist, then it performs no repartition, and serves the request remotely, incurring the cost $1$.
    Hence, in the following we bound the cost of moving to component-respecting partition.

@@

  \begin{theorem}
    \TAlg is $O(\ell)$-competitive.
  \end{theorem}
  \begin{proof}
    Fix a phase of \TAlg, and consider the state of its counters at the end of it.

    Let $A^I$ be the cost of (extra-cluster) communication incurred in this phase by \TAlg between pairs that belonged to the single component at the end of the phase.
    Let $A^E$ be the cost of (extra-cluster) communication incurred in this phase between the nodes that belong to different components throughout this phase.
    Let $A^M$ be the cost of migrations performed by \TAlg in this phase.

    We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.

    During a single phase, at most $2\cdot \ell$ component merge operations are performed.
    Exceeding this number would mean that a component of size $4$ would emerge, a contradiction.
    Together with Lemma~\ref{lem:1req}, this gives us $A^M \leq 6\cdot\alpha\cdot\ell$.

    Now we bound $A^I$.
    Each component of size $3$ contributes at most $(3\cdot \alpha - 1)$ to $A^I$, as at most $2$ of its pairs are saturated and contribute $\alpha$ each, and the unsaturated pair contributes at most $\alpha-1$.
    Each of $\ell$ clusters of \TAlg contain at most one component of size $3$ (other cluster configurations contribute less: $3$ singletons contribute $0$ and a component of size $2$ contributes at most $\alpha$ to $A^I$), and in total, $A^I \leq \ell \cdot (3\cdot \alpha-1)$.

    We bound $A^E$ by $k^2 \cdot (\alpha - 1)$, as no more than $k^2$ pairs are unsaturated, and each of them contributes at most $\alpha -1$.

    In total, the cost of \TAlg is at most $A^I + A^E + A^M$ during the phase.

    Now we lower bound the cost of $\OPT$.
  \end{proof}
